---
title: "Homework 9"
subtitle: Managerial Statistics
author: "Leticia Liakopulos"
output:
  html_document: default
  html_notebook: default
---

This R Notebook file *homework9.Rmd* contains the questions and 
the R chunks to answer them. 

The data (stored in mtcars491.csv) was extracted from the Motor Trend 
US magazine, and comprise fuel consumption and 10 aspects of automobile
design and performance for 32 automobiles.

#### Variables:
mpg	  : Miles/(US) gallon    
cyl	  : Number of cylinders     
disp  :	Displacement (cu.in.). Overall volume in the engine.     
hp	  : Gross horsepower     
drat  :	Rear axle ratio      
wt	  : Weight (1000 lbs)       
qsec  :	1/4 mile time. Fastest time to travel 1/4 miles from standstill.    
engine: Engine type  (v = V-shaped, s = straight)     
trans : Transmission (a = automatic,  m = manual)      
gear  :	Number of forward gears (three=1, four=4, five=5)     
carb  :	Number of carburetors      

#### Initial Work
0. Download the homework9A.Rmd and mtcars491.csv files 
   from the course website.     

Read data from mtcars491.csv and assign it to the object car.
```{r}
mtcars <- read.csv("mtcars491.csv")
attach(mtcars)
```

#### Question 1. Introduction and Descriptive Statistics		  CLASS 1      

What is the mean mpg of the cars in this data?
```{r}
mean(mpg)
```
(b)	Calculate the mean mpg's of the cars with automatic and manual
    transmissions separately.
```{r}
tapply(mtcars$mpg, mtcars$trans, mean)
```

(c)	What proportion of the cars in this data have automatic transmissions?
```{r}
prop.table(table(mtcars$trans))
```
**Answer**    

(d) Boxplots of mpg for the cars with automatic and manual transmissions.
```{r}
boxplot(mpg~trans, data=mtcars, col="red", names=c("automatic","manual"))
```

(e)	Display the histogram of the mpg's.
```{r}
hist(mpg, col = "orange")
```

#### Question 2. Normal Distribution and Central Limit Theorem	  CLASS 2     
The EPA states that the city mpg's of Toyota Prius are normally distributed with a mean (mu) of 30 and a standard deviation (sigma) of 4 miles.

What is the probability that 	    				
(a)	a randomly selected Prius will have a city mpg less than 34 miles?
```{r}
pnorm(34,30,4)
```

(b) a randomly selected Prius will have a city mpg between 26 and 34 miles?
```{r}
pnorm(34,30,4)-pnorm(26,30,4)
```

(c) a randomly selected Prius will have a city mpg higher than 36 miles?
```{r}
pnorm(36,30,4)
```

(d) Suppose now that 16 cars (Prius) are randomly selected. What is the
    probability that their mean city mpg will be less than 28 miles?				
```{r}
pnorm(28,30,1)
```

#### Question 3. Interval Estimation		       CLASS 3         
(a)	Based on the sample results, What is the point estimate 
    for the mean mpg (mu) of all cars?	
```{r}
mean(mtcars$mpg)
```

(b)	Based on the sample results, determine 95% confidence
interval estimate for the mean mpg (mu) all cars.
```{r}
t.test(mtcars$mpg)
```
**Answer:**  95% CIE= LL= 17.91768 UL= 22.26357         


#### Question 4. Hypothesis Testing (One-Sample Test)		    CLASS 4         
(a)	The past historical data showed that the mean mpg of the cars was 
    18 miles. Due to the improvements in the design technology, it is
    claimed that the mean mpg has increased. Does the sample data
    provide a strong evidence to indicate that the mean GPA has
    increased at a significance level (alpha) of 0.05?
    
```{r}
options(scipen=999)
```


(i) State the hypotheses:        
**Answer**    
Hypotheses:     
Ho: x > mu                   
HA: x <= mu                  

Calculate the test statistic and critical value.
```{r}
t.test(mtcars$mpg, mu=18, alt ="greater", conf.level=0.95)
```

(ii) Decision Rule:     
**Answer**    
Reject Ho if p_value < 0.05

(iii) Decision:     
**Answer**        
p= 0.02938 < 0.05
Reject Null.

(iv) Conclusion:     
**Answer**   
The sample provides enough evidence that the mu of mpg has increased.

#### Question 6. Two-Sample and Multiple Sample Tests	 	  CLASS 6     
(a)	Does the sample show if there is significant difference between 
the mean mpg's of cars with automatic and manual transmissions at a significance level of 0.05?

(i) State the hypotheses:          
**Answer**      
Hypotheses:    
Ho: muA = muB       
HA: muA  ≠ muB      

Calculate the test statistic and critical value.
```{r}
t.test(mpg~trans, data=mtcars)
```
(ii) Decision Rule:     
**Answer**    
Reject Ho if p-value < a=0.05
Do not reject Ho if p-value > a=0.05 

(iii) Decision:       
**Answer**    
p-value= 0.001374 < a= 0.05

(iv) Conclusion:    
**Answer**
Reject Ho. The mean mpg for manual and automatic transmissions are not the same.

   
(b)	Does the sample show if the the mean mpg's of the car with 
    different number of gears are the same (or not the same) at
    a significance level of 0.05?

(i) State the hypotheses:    
**Answer**   
Hypotheses:
Ho: muA = muB= muC    
HA: Not all the same ( at least one is different) 

Calculate the test statistic and critical value.
```{r}
TestGEAR <- aov(mpg~gear, data=mtcars)
summary(TestGEAR)
```

(ii) Decision Rule:    
**Answer**    
Reject Ho if p-value < a=0.05
Do not reject Ho if p-value > a=0.05   

(iii) Decision:      
**Answer**    
p=0.000295 < a=0.05

(iv) Conclusion:      
**Answer**
Do not reject Ho. The mpg for the cars with different number of gears is not 
the same.
   

(c) If the mean mpg's are different, which number of gears have 
have a different mean mpg from each other at a significance level of 0.05?
```{r}
TukeyHSD(TestGEAR)
```
**Answer**     
The number of gears that have a different mpg at a significance level of 0.05 are three and four.

#### Question 8. Regression		         CLASS 7 and 8         
You want to predict the mpgs by using a regression model.

(a)	Using R, obtain the multiple regression model (M2) to predict 
    the mpg's of the cars by using the following 2 variables 
    (qsec=1/4 mile time and trans=transmission)
```{r}
R2<- lm(mpg~ qsec+trans, data=mtcars)
```

(b)	Display the summary output of the regression model (R6).
```{r}
summary(R2)
```

(c) What is the estimated regression equation?    
**Answer** 
y= -18.8893 + 1.9819X1 + 8.8763X2


(d) What does the coefficient of trans (b2) tell us?     
    i.e., What is the interpretation of b2?    
**Answer** 
The coefficient of trans(b2)=8.8763 indicates the change in the mpg of a car with automatic and manual transmition.


(e) Which independent variables are not significant in this
    regression at a significance level of 0.05?        
**Answer**  
All variables are significant since all of their p-values are lower than 0.05.
  

(f) What percentage of the total variation is explained 
    by this regression?        
**Answer**   
68.68% of the total variation of the mpg is explained in the regression.


#### Question 9. Stepwise Regression     CLASS 9     
(a)	Use the backward elimination method to determine the
    explanatory (x) variables to be used in a final model.
```{r}
fitAll <- lm(mpg~., data=mtcars)
summary(fitAll)
step(fitAll, direction = "backward")
```

(b1)	Which explanatory variables are to be included in the final model?   
**Answer**    
wt, qsec, trans

(b2) Obtain the regression model using the variables selected
    by the stepwise regression.
```{r}
M3<- lm(mpg~ wt+ qsec+ trans, data=mtcars)
summary(M3)
```

(c) What percentage of the total variation is explained by 
    this regression?      
**Answer**
84.97% of the total variation is explained on this regression.



(d)	Using this regression model obtained above in (b2), 
    calculate 95% confidence interval estimate (CIE) for 
    the mean mpg of all cars with the following values:    
 mpg cyl disp  hp  drat   wt  qsec engine trans gear carb    
21.0	6	 230  140	 3.60  3.20 18.0   v      m   four   3    
```{r}
predict(M3 , data.frame(mpg=21.0, cyl=6, disp=230, hp=140, drat=3.60, wt=3.20, qsec=18.0, engine="v", trans="m", gear="four", carb=3) ,interval = "confidence")
```

## After completing your R codes, create an html file 
that will contain       
   -	all the codes       
   -	executed outputs and      
   -	answers the follow-up questions, if there is any.    

Complete the following steps to create the html file:    
Step 1. Execute each code chunk one by one (starting from the first chunk)    
Step 2. Save the file (Save As…). Use file name:  homework9    

If you have completed Step 1 correctly, then Step 2 will save your Rmd file and (at the same time) will create an html file on your working directory. 

Submit this html file as your homework9 on Sakai.

